---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

# Euphausia pacifica

## Model
```{r}
Mel6 <- lmer(length ~ sex*temp_2_z + temp_100_z + sex:temp_100_z + sst_sd_z + chla_z + beuti_z + sla_z + sex:sla_z + moci_spring_z + sex:moci_spring_z + (1|station), data = ep)

summary(Mel6)
sjPlot::plot_model(Mel6)
```

## Fixed Effects

> SST stands out as a particularly important variable with a positive impact on overall EP size. However, this appears to be mostly driven by Southern California.

```{r}
epg <- summarize(group_by_at(ep, vars(sites, year, region, temp_2_z, sla_z, sst_sd_z)), length = mean(length))
ggplot(epg, aes(x = temp_2_z, y = length)) + 
  geom_point(aes(color = sites)) + 
  geom_smooth(method = 'lm', formula = y ~ splines::bs(x, 2), se = T)
```

> When Southern CA (San Diego) is removed from the data, the relationship is still positive, but looks more asymptotic. This is consistent with the N vs S change in temp~length relationship. 

```{r}
ggplot(filter(epg, sites != "SouthernCA"), aes(x = temp_2_z, y = length)) + 
  geom_point(aes(color = sites)) + 
  geom_smooth(method = "lm", formula = y ~ splines::bs(x, 2), se = T)
```


```{r}
ggplot(epg, aes(x = sst_sd_z, y = length)) + 
  geom_point(aes(color = sites)) + 
  geom_smooth(method = "lm")
```


> SST_SD and MOCI had a positive effect on size, but weaker than SST. BEUTI, SLA, and Chl-a had a weak negative impact on body size. There is also some indication that males and females responded differently to temperature and SLA fluctuations.

## Random Effects

> Sites in the north have larger individuals than sites in the south. Onshore EP tend to be larger than offshore counterparts. San Diego offshore are very very small. 

```{r}
groups <- data.frame(
  station = as.numeric(row.names(ranef(Mel6)$station)),
  intercept = ranef(Mel6)$station$'(Intercept)'
)
groups <- left_join(groups, regions)
ggplot(groups) +
  geom_point(aes(x = latitude, y = intercept, color = sites, shape = shore), size = 3) + 
  geom_hline(aes(yintercept = 0))
```


# Thysanoessa spinifera

## Model

```{r}
Mtl2 <- lmer(length ~ sex*temp_2_z + temp_100_z + sex:temp_100_z + sst_sd + chla_z + beuti_z + moci_spring_z + cuti_z + (1|station), data = ts)

summary(Mtl2)
sjPlot::plot_model(Mtl2)
```

## Fixed Effects

